---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I Learned on Jan. 27

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```


### Summarizing

I need to use group_by and summarize. Here's an example of grouping by county and calculating counts, sum and other descriptive statistics.


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```
9/26
Mutate to clean up data
Fix a wrong value in a column and change the data type
e.g.:
```{r}
texas_precinct_20 <- texas_precinct_20 %>%
  mutate(election_day = case_when(
    election_day == '5+' ~ '49',
    TRUE ~ election_day
  ))

texas_precinct_20 <- texas_precinct_20 %>% mutate(election_day = as.numeric(election_day))
```

Used mutate and case_when to replace incorrect 5+ with 49 -- don't forget TRUE to keep the other values as they were
Used mutate and as.numeric to make election_day values into numbers


clean_names() from "janitor" library standardizes column headers to eliminate all caps, add underscores instead of spaces, make sure none of them starts with numbers...

```{r}
cleaned_conowingo <- conowingo %>%
  clean_names()
```

rename() allows you to rename a column header
```{r}
cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber)
```

distinct() eliminates duplicates
get_dupes() shows you the duplicates


str_sub to make all zip codes 5 numbers long
```{r}
# cleaning function
cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
```

str_to_title() to standardize capitalization in, for instance, city names
```{r}
# cleaning function
cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city))
```

str_detect with the ^ says, "Look in the city column. If you find something that starts with Conowing, change it to Conowingo."
```{r}
# cleaning function
cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L)) %>%
  mutate(city = str_to_title(city)) %>%
  mutate(city = case_when(
    str_detect(city,"^Conowing") ~ "Conowingo",
    TRUE ~ city
  ))
  

### Packages

Lubridate - makes it easier to work with dates, times and timespans; gets R to recognize that you have the dates and times you think you do
