---
title: "pre_lab_14.Rmd"
author: "derek willis"
date: "12/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# An intro to text analysis

Throughout this course, we've been focused on finding information in structured data. We've learned a lot of techniques to do that, and we've learned how the creative mixing and matching of those skills can find new insights.

What happens when the insights are in unstructured data? Like a block of text?

Turning unstructured text into data to analyze is a whole course in and of itself -- and one worth taking if you've got the credit hours -- but some simple stuff is in the grasp of basic data analysis.

To do this, we'll need a new library -- [tidytext](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html), which you can guess by the name plays very nicely with the tidyverse. So install it in the console with `install.packages("tidytext")` and we'll get rolling.

### Task 1: Load libraries and settings
**Task** Run the following code in the gray-colored codeblock below to load the libraries we'll use.

```{r}
#install.packages("tidytext")
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
```

Remember the end of the previous lab, where you scraped all of those press release links from Maryland Sen. Ben Cardin's website? We're going to take that one step further and analyze the actual text of those releases using tidytext. Our starting question: what words or phrases appear most in Cardin's press releases?

To answer this question, we'll use the text of those releases. For the scraping lab we gathered the URLs of 999 of Cardin's press releases, and for this exercise we'll be working with the text of them. 

Let's read in this data and examine it:

### Task 2: Read in data
**Task** Run the following code and describe the dataframe it outputs.
**Answer** The dataframe has the date, title, url and text of Cardin's 999 press releases.

```{r}
releases <- read_rds("data/cardin_releases.rds")
```

As an aside, below is an example of how you would scrape the text from the first 10 press releases. We already have the text in the dataframe, but this lets you know how it was collected.

### Task 3: Example of gathering text
**Task** Run the following code to gather the text for the first 10 press releases.

```{r}

urls <- releases %>% top_n(10) %>% pull(url)

release_text <- tibble(url = character(), text = character())

# loop over each url in the list of urls
for (u in urls){
  # wait a fraction of a second so we don't hammer the server
  Sys.sleep(0.2)
  # read in the html from the url
  html <- u %>% read_html()
  # use the xpath of the text of the release to grab it and call html_text() on it
  text <- html %>% 
    html_element(xpath="/html/body/div/div/div/div/div/div/div[2]/div[1]/div/div[4]") %>% 
    html_text()
  
  release_text <- release_text %>% add_row(url = u, text = str_squish(text))
}

release_text
```

What we want to do is to make the `text` column easier to analyze. Let's say we want to find out the most commonly used words. We'll want to remove URLs from the text of the releases since they aren't actual words. Let's use mutate to make that happen:

### Task 4: Remove URLs from content
**Task** Run the following code.

```{r}
releases <- releases %>%
  mutate(text = gsub("http.*","", text))
```

If you are trying to create a list of unique words, R will treat differences in capitalization as unique and also will include punctuation by default, even using its `unique` function:

### Task 5: Trying out unique
**Task** Run the following code and describe what the `unique` function does to the original list of words.
**Answer**
It lists the separate capitalizations of dog and cat as separate words, and it includes item in the list that is just a comma as its own word. But for items that are identical, like "dog" and "dog," it only lists dog once.
```{r}
a_list_of_words <- c("Dog", "dog", "dog", "cat", "cat", ",")
unique(a_list_of_words)
```

Fortunately, this is a solved problem with tidytext, which has a function called `unnest_tokens` that will convert the text to lowercase and remove all punctuation. The way that `unnest_tokens` works is that we tell it what we want to call the field we're creating with this breaking apart, then we tell it what we're breaking apart -- what field has all the text in it. For us, that's the `text` column:

### Task 6: Trying out unnest_tokens
**Task** Run the following code and describe the output of using the `unnest_tokens` function.
**Answer**
The result is a column called "unique_words" containing all 603,725 words in Cardin's press releases. The words are all lowercased, and any ending punctuation has been removed, although punctuation inside a word (like in 'u.s') remains. The column also includes single letters such as d.
```{r}
unique_words <- releases %>% select(text) %>%
  unnest_tokens(word, text)
View(unique_words)
```

### Task 7: Make a column chart
**Task** Run the following code and describe what the resulting graphic shows. Is it interesting?
**Answer** Unfortunately, it's pretty boring. It almost exclusively contains common articles and conunctions like "a," "the" or "and." And also Maryland!

Now we can look at the top words in this dataset. Let's limit ourselves to making a plot of the top 25 words, and we'll use the function `count` to do the counting:

```{r}
unique_words %>%
  count(word, sort = TRUE) %>%
  top_n(25) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
      labs(x = "Count",
      y = "Unique words",
      title = "Count of unique words found in Cardin releases")
```

Well, that's a bit underwhelming - a lot of very common (and short) words. This also is a solved problem in working with text data, and words like "a" and "the" are known as "stop words". In most cases you'll want to remove them from your analysis since they are so common. Tidytext provides a dataframe of them that we'll load, and then we'll add some of our own.

### Task 8: Load the stop words
**Task** Run the following code

```{r}
data("stop_words")

stop_words <- stop_words %>% 
  add_row(word = "ben") %>% 
  add_row(word = "cardin") %>% 
  add_row(word = "senator") %>%
  add_row(word = "senators") %>%
  add_row(word = "maryland") %>%
  add_row(word = 'federal') %>% 
  add_row(word = 'u.s') %>% 
  add_row(word = 'md') %>% 
  add_row(word = 'senate') %>% 
  add_row(word = "hollen") %>% 
  add_row(word = "van") %>% 
  add_row(word = "chris")

```

Then we're going to use a function we haven't used yet called an `anti_join`, which filters out any matches. So we'll `anti_join` the stop words and get a list of words that aren't stop words.

From there, we can get a simple word frequency by just grouping them together and counting them. We can borrow the percent code from above to get a percent of the words our top 10 words represent.

### Task 9: Using anti_join
**Task** Run the following code and describe the results. Is it more interesting than before?
**Answer** This is more interesting! Some of the words seem to be pretty boilerplate for a public official: "community" and "communities," "public," "program," "act"...."Health" is the top term, probably partly because of covid, but maybe Cardin focuses on health issues anyway. "Covid" is number 10, which is lower than I would have expected.

```{r}
unique_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

Those seem like more relevant unique words. Now, here's where we can start to do more interesting and meaningful analysis. Let's create two dataframes of unique words based on time: one for all of 2021 and the other for all of 2022:

### Task 10: Create dataframes for 2021 and 2022
**Task** Run the following code

```{r}
unique_words_2021 <- releases %>%
  filter(date < '2022-01-01') %>%
  select(text) %>%
  unnest_tokens(word, text)

unique_words_2022 <- releases %>%
  filter(date >= '2022-01-01') %>%
  select(text) %>%
  unnest_tokens(word, text)
```

Then we can create top 10 lists for both of them and compare:

### Task 11: Create dataframes with the top 10 words in each year
**Task** Run the following code and describe the results.
**Answer** "Health," "covid," "19," and "pandemic" all made the top 10 in 2021, which makes sense, but only "health" remained in 2022, as the political world and public discourse moved away from covid. "Baltimore" appeared in the top 10 in 2022. Was Baltimore in Cardin's top 10 before the pandemic?

```{r}
unique_words_2021 %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)

unique_words_2022 %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

In the 2021 top 10 list, "covid" and "pandemic" appear, which makes sense, while the 2022 doesn't reference the same topic aside from the more generic "health".

## Going beyond a single word

The next step in text analysis is using `ngrams`. An `ngram` is any combination of words that you specify. Two word ngrams are called bigrams (bi-grams). Three would be trigrams. And so forth.

The code to make ngrams is similar to what we did above, but involves some more twists.

So this block is is going to do the following:

1. Use the releases data we created above, and filter for pre-2022 releases.
2. Unnest the tokens again, but instead we're going to create a field called bigram, break apart summary, but we're going to specify the tokens in this case are ngrams of 2.
3. We're going to make things easier to read and split bigrams into word1 and word2.
4. We're going to filter out stopwords again, but this time we're going to do it in both word1 and word2 using a slightly different filtering method.
5. Because of some weirdness in calculating the percentage, we're going to put bigram back together again, now that the stop words are gone.
6. We'll then group by, count and create a percent just like we did above.
7. We'll then use top_n to give us the top 10 bigrams.

### Task 12: Create a dataframe with the top 10 two-word phrases for 2021
**Task** Run the following code and describe the results.
**Answer**
The results suggest that covid 19 and Biden's American Rescue Plan, as well as health care and public health, occupied prominent places in Cardin's 2021 press releases. The Chesapeake Bay seems like a bit of an outlier, if probably a safe thing for a politician to talk about in Maryland.  But I wonder what policy on the Chesapeake Bay he may have been referring to. Also, I wonder what in particular he was saying or doing about human rights. Human rights as they relate to the pandemic?
```{r}
releases %>%
  filter(date < '2022-01-01') %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  mutate(bigram = paste(word1, word2, sep=" ")) %>%
  group_by(bigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

And we already have a different, more nuanced result. Health was among the top single words, and we can see that "health care", "human rights" and "chesapeake bay" are among the top 2-word phrases. What about after 2021?

### Task 13: Create a dataframe with the top 10 two-word phrases for 2022
**Task** Run the following code and describe the results.
**Answer** "Covid 19," "health care" and "human rights" are all still in the top ten. "Mental health" entered the top 10 in 2022. There are also more place names - Baltimore and Prince George's in addition to the Chesapeake. Is "project location" in reference to a particular project?

```{r}
releases %>%
  filter(date >= '2022-01-01') %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  mutate(bigram = paste(word1, word2, sep=" ")) %>%
  group_by(bigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

While "covid 19" is still a top phrase, it's not _the_ leading phrase any longer. "Mental health" makes an appearance, too. You'll notice that the percentages are very small; that's not irrelevant but in some cases it's the differences in patterns that's more important.

There are some potential challenges to doing an analysis. For one, there are variations of words that could probably be standardized - maybe using OpenRefine - that would give us cleaner results. There might be some words among our list of stop words that actually are meaningful in this context.

## Sentiment Analysis

Another popular use of text analysis is to measure the sentiment of a word - whether it expresses a positive or negative idea - and tidytext has built-in tools to make that possible. We use word counts like we've already calculated and bring in a dataframe of words (called a lexicon) along with their sentiments using a function called `get_sentiments`. The most common dataframe is called "bing" which has nothing to do with the Microsoft search engine. Let's load it:

### Task 14: Load the bing lexicon and produce sentiments for our 2021 and 2022 unique words
**Task** Run the following code and describe the results. Do any of the sentiments seem incorrect or counter-intuitive?
**Answer** The idea that "work" is necessarily positive probably says something about our society's relationship to capitalism (Max Weber's "Protestant Work Ethic and the Spirit of Capitalism"). Also, "lead" may be positive, but what about "lead" poisoning? And "risk" can be positive. This seems like a very conservative sentiment analysis.

```{r}
bing <- get_sentiments("bing")

bing_word_counts_2021 <- unique_words_2021 %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE)

bing_word_counts_2022 <- unique_words_2022 %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE)

View(bing_word_counts_2021)
View(bing_word_counts_2022)
```

Gauging the sentiment of a word can be heavily dependent on the context, and as with other types of text analysis sometimes larger patterns are more meaningful than individual results. But the potential with text analysis is vast: knowing what words and phrases that public officials employ can be a way to evaluate their priorities, cohesiveness and tactics for persuading voters and their colleagues. And those words and phrases are data.
