---
title: "lab_01"
author: "Derek Willis"
date: "7/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
wv_winred_contribs <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** 13,382 rows and 24 columns

```{r}
glimpse(wv_winred_contribs)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?
**A2** Martinsburg

```{r}
group_by(city) %>% 
  summarise(count_contribs = n()) %>% 
  arrange(desc(count_contribs))
```
```{r}
group_by(city) %>% 
    +   summarise(count_contribs = n(),) %>% 
    +   arrange(desc(count_contribs))
```
```{r}
wv_winred_contribs
group_by(city) %>% 
    +   summarise(count_contribs = n(),) %>% 
    +   arrange(desc(count_contribs))
```
```{r}
wv_winred_contribs %>% 
  group_by(city) %>% 
    +   summarise(count_contribs = n()) %>% 
    +   arrange(desc(count_contribs))
```  
```{r}
wv_winred_contribs %>% 
  group_by(city) %>% 
  summarise(count_contribs = n()) %>% 
  arrange(desc(count_contribs))
```

  I have no idea why that worked but the one I did immediately before didn't.
  
  
**Q3** Which West Virginia city accounted for the fewest contributions? How many?
**A3** There are 40 cities that had only one contribution.

```{r}
wv_winred_contribs %>% 
  group_by(city) %>% 
  summarise(count_contribs = n()) %>% 
  arrange(count_contribs)
```
**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** Martinsburg is one of the larger cities in WV, so there are more people who could make campaign contributions. Martinsburg is also located in the eastern panhandle of the state, which is closer to Washington, D.C. - some people even commute from there - and is known for being demographically distinct from other parts of the state. People might be wealthier and have more expendable income. I'm also guessing it's a more blue-leaning part of the state. Maybe, in a heavily Republican state, people with more expendable income are trying to use their wealth to support Democratic candidates and change the political hue of the state.

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?
**A5** 
Save America Joint Fundraising Committee got the largest total amount of money ($66,972.89), with 2,448 donations. The average donation amount was $27.36.
```{r}
wv_winred_contribs %>% 
  group_by(memo_text) %>% 
  summarise(count_contribs = n(),
            total_amount = sum(amount),
            mean_amount = mean(amount)) %>% 
  arrange(desc(total_amount))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** The Save America Joint Fundraising Committee is former President Donald Trump's political action committee. The people who made donations might be hoping to see Trump run for president.

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by city):
* count of contributions
* total amount of all contributions
* the average contribution amount
* the median contribution amount
* the smallest contribution amount
* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7**. Most contributions are modest, with a relatively small number of large contributions. I can see this because the mean is $26.70 and the median is even smaller, at $10. I wonder who's making contributions of 1 cent - the minimum amount - and why. Also, if Martinsburg has the most contributions, as we saw earlier in the lab, I wonder if those contributions tend to be large or small. 

```{r}
wv_winred_contribs %>% 
  summarise(count_contribs = n(),
  total_amount = sum(amount),
  mean_amount = mean(amount),
  median_amount = median(amount)) %>% 
  arrange(desc(amount))
  
```


```{r}
wv_winred_contribs %>% 
  group_by(state, amount) %>% 
  summarise(count_contribs = n(),
  total_amount = sum(amount),
  mean_amount = mean(amount),
  median_amount = median(amount)) %>% 
  arrange(desc(amount))

```

```{r}
wv_winred_contribs %>%
  group_by(state) %>% 
  summarise(count_contribs = n(),
            total_amount = sum(amount),
            mean_amount = mean(amount),
            median_amount = median(amount),
            min_amount = min(amount),
            max_amount = max(amount)) %>%  
  arrange(desc(amount))
```
```{r}
wv_winred_contribs %>% 
  group_by(state) %>%
  summarise (count_contribs = n(),
             total_amount = sum(amount),
             mean_amount = mean(amount),
             median_amount = median(amount),
             min_amount = min(amount),
             max_amount = max(amount)) %>%
  arrange((total_amount))
```
```{r}
wv_summary <- wv_winred_contribs %>% 
  group_by(state) %>%
  summarise (count_contribs = n(),
             total_amount = sum(amount),
             mean_amount = mean(amount),
             median_amount = median(amount),
             min_amount = min(amount),
             max_amount = max(amount)) %>%
  arrange((total_amount))
```

