---
title: "lab_04"
author: "derek willis"
date: "8/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Load the libraries you will need to read in and clean up the data, knowing that you can go back and add a library and re-run this block after you get a look at the dataframes.
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll use Maryland 2022 primary precinct election results to do this, and our main question will be this: Which precincts had the most votes cast at them statewide, counting both the Republican and Democratic primaries? To answer this question, we'll need to first combine the two parties' primary precinct results into a single dataframe and then join it to data that has more information about each voting location.

First, load the two precinct results CSV files in the `data` folder, one each for the Democratic and Republican primaries, and then combine them into a single dataframe using the method we used in the pre_lab. You'll need to clean up the column names to make the data easier to work with - make sure you have loaded the library to do that.

```{r}
md_primary_precinct_dem <- read_csv("data/md_primary_precinct_dem.csv")
md_primary_precinct_rep <- read_csv("data/md_primary_precinct_rep.csv")

glimpse(md_primary_precinct_dem)
glimpse(md_primary_precinct_rep)

md_primary_precinct_combined <- bind_rows(list(md_primary_precinct_dem, md_primary_precinct_rep))
View(md_primary_precinct_combined)
```
```{r}
cleaned_md_primary_precinct_combined <- md_primary_precinct_combined %>%
  clean_names()

cleaned_md_primary_precinct_combined
```
```{r}
cleaned_md_primary_precinct_combined <- cleaned_md_primary_precinct_combined %>% mutate(election_district_precinct = as.character(election_district_precinct))
```


Next, let's read in the precinct voting location data and make sure you clean up the column names:

```{r}
precinct_voting_locations <- read_csv("data/precinct_voting_locations.csv")
```
```{r}
cleaned_precinct_voting_locations <- precinct_voting_locations %>%
  clean_names()

cleaned_precinct_voting_locations
```

These two dataframes have a pair of columns in common - not necessarily with the same column names. Join the two dataframes on that unique combination that describes a precinct and the jurisdiction that contains it:

```{r}
md_precinct_results_with_locations <- cleaned_md_primary_precinct_combined %>% 
  left_join(cleaned_precinct_voting_locations, by=c("election_district_precinct"="precinct_code",
                                                    "county"="county_code"))
md_precinct_results_with_locations
```

## Answer questions

Q1. Using the newly-created dataframe and isolating just the results for Wes Moore and Aruna Miller, write code that calculates how many voting locations each county has, showing the county with the most voting locations first. There's a catch: multiple precincts can share the same voting location, so you'll need to calculate the distinct number of locations (Google is your friend).

Does the county with the most voting locations have the largest population in the state (ask yourself: where have you seen county population figures recently)?

A1. No. Prince George's County has the most voting locations (240) and the second largest population in the state (967,201). Montgomery County has the largest population (1,062,061) but only the second most voting locations (225).

```{r}
Moore_Miller_precinct_results <- md_precinct_results_with_locations %>% 
  filter(candidate_name == "Wes Moore and Aruna Miller")

Moore_Miller_precinct_results
```

```{r}
Moore_Miller_precinct_results %>% 
  group_by(county_name) %>% 
  summarize(distinct_locations = n_distinct(location_name)) %>% 
  arrange(desc(distinct_locations))
```


Q2. Write code to produce a `votes` column which is the result of adding together `early_votes`, `election_night_votes`, `provisional_votes` and the mail-in votes, overwriting your dataframe so that it now has that `total_votes` column going forward. Then, separately write code to show each county and precinct and calculate the total votes in the Governor's race - bearing in mind that each precinct will have multiple rows in the dataframe, but you only want to see one row for each precinct. What is the name and voting location of the precinct that saw the most votes in the governor's race? 

A2. Dr. Henry A. Wise, Jr. High School in precinct 15005 had the most votes in the governor's race (2,396).

```{r}
md_precinct_results_with_locations <- md_precinct_results_with_locations %>% 
  mutate(votes = early_votes + election_night_votes + provisional_votes + mail_in_ballot_1_votes + mail_in_ballot_2_votes)
```
```{r}
md_precinct_results_with_locations %>%
  filter(office_name == "Governor / Lt. Governor") %>% 
  group_by(county_name, election_district_precinct) %>% 
  summarize(total_votes = sum(votes)) %>% 
  arrange(desc(total_votes))
```

```{r}
md_precinct_results_with_locations %>%
  filter(office_name == "Governor / Lt. Governor") %>% 
  filter(election_district_precinct == "15005" & county_name == "Prince George's County")
```
```{r}
md_precinct_results_with_locations %>%
  filter(office_name == "Governor / Lt. Governor") %>% 
  filter(election_district_precinct == "15005" & county_name == "Prince George's County") %>%
  summarize(total_votes = sum(votes))
```
Q3. Let's look at the average number of votes in each party's primary race for governor. Take the code you wrote for the last part of Q2, save the results to a new variable and then find the precincts with 0 total votes for governor. How many precincts are there, and which county has the most? What else would you need to know about these precincts to know whether this might be worth a story?

Finally, write code to calculate the average and median number of total votes across all precincts statewide (so you want to see just two numbers - the average and median). To do this, you will need to use the `ungroup()` function on your new dataframe _before_ you do the math - https://www.rdocumentation.org/packages/multiplyr/versions/0.1.1/topics/ungroup. There's no state column here, so you aren't grouping by anything, but can still use summarize to do the math.

Take a look at the the Annotated Code of Maryland's election laws (https://mgaleg.maryland.gov/2023RS/Statute_Web/gel/gel.pdf) - can you find a limit on the number of voters assigned to a precinct? How might your answer impact wait times to vote?

A3. There are 12 precincts where no votes were cast in the Governor/Lt. Governor race, and Prince George's County has more of these precincts than any other county (4). In order to decide if this merits a story, I would want to know how voters in these four precincts voted in other races. How many voters were assigned to these precincts, and how many of them voted in any race? What are these four precincts like demographically, and are they demographically similar to each other? Was their 2022 voting substantially different from other years?

The average number of votes cast across precincts statewide was 466 (to the nearest vote). The median was 418.

I can't find a limit on the number of voters assigned to a precinct in Maryland's election laws. No upward limit on voters per precinct could cause very long wait times in crowded precincts. This makes me wonder how often precincts are rethought. Are there adjustments if one precinct becomes really densely populated, for instance?


```{r}
md_governor_votes <- md_precinct_results_with_locations %>% 
  filter(office_name == "Governor / Lt. Governor") %>% 
  group_by(county_name, election_district_precinct) %>% 
  summarize(total_votes = sum(votes))

md_governor_votes %>% 
  arrange(total_votes)
```


```{r}
md_governor_votes %>%
  ungroup() %>% 
  summarize(mean_votes = mean(total_votes),
            median_votes = median(total_votes))
```


Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve candidates, or counties or types of voting, but it cannot be a variation of a previous question (so, for example, you can't ask, "What's the average number of votes per precinct in the U.S. Senate primary?"). First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me or Ramsay for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.

A4: The question I started with was: How did voters in the four Prince George's County precincts with no gubernatorial votes vote in other races? I chose to focus on precinct 13029.

First, I filtered to see only results for precinct 13029, which showed me that the precinct straddles multiple counties. Then I filtered for precinct 13029 and Prince George's County. It actually looks to me like no one cast votes in any race in the part of 13029 in Prince George's County.

So then I wondered how many votes were cast in parts of 13029 located in other counties. I noticed before that parts of precinct 13029 are in Montgomery County, so next I filtered for precinct 13029 and Montgomery County. These voters seem to have been assigned to Highland Elementary School in Silver Spring, and 676 people cast votes in the gubernatorial race there.

How did people who voted at Highland Elementary School vote in other races? Fewer people (668) voted in the race for U.S. Senator than in the gubernatorial race. And 679 voted for County Executive.

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029" )
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029",
         county_name == "Prince George's County")
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029" & county_name == "Montgomery County")
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029" & county_name == "Montgomery County" & office_name == "Governor / Lt. Governor")
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029" & county_name == "Montgomery County" & office_name == "Governor / Lt. Governor") %>%
  summarize(total_votes = sum(votes))
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029") %>% 
  filter(county_name == "Montgomery County") %>% 
  filter(office_name == "U.S. Senator") %>% 
  summarize(total_votes = sum(votes))
```

```{r}
md_precinct_results_with_locations %>% 
  filter(election_district_precinct == "13029") %>% 
  filter(county_name == "Montgomery County") %>% 
  filter(office_name == "County Executive") %>% 
  summarize(total_votes = sum(votes))
```